---
title: "Trying to use half-hourly NEE data for the fVAR algorithm"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r}
devtools::load_all(".")
library(rsofun)
library(dplyr)
library(readr)
library(lubridate)
library(ggplot2)
```


# Get data

```{r}
usesite <- "CH-Lae"

getvars <- c( 
  "GPP_NT_VUT_REF", 
  "GPP_DT_VUT_REF",                
  "LE_F_MDS", 
  "LE_F_MDS_QC", 
  "NEE_VUT_REF", 
  "NEE_VUT_REF_QC", 
  "TA_F",
  "VPD_F",
  "P_F",
  "SW_IN_F",
  "NETRAD"
)

df_fluxnet <- get_obs_bysite_fluxnet2015( 
  sitename=usesite, 
  path_fluxnet2015 ="~/data/FLUXNET-2015_Tier1/20191024/HH/", 
  timescale="hh", 
  getvars=getvars,
  threshold_GPP = 0,      # take only measured data
  getswc = TRUE
  ) %>% 
  mutate(sitename = usesite) %>% 
  
  ## use daytime data only
  dplyr::filter(ppfd > 0.0) %>% 
  
  ## use growing season of CH-Lae only
  dplyr::filter(yday(date)>121 & yday(date)<121+161) %>% 
  
  ## additional cleaning (only positive VPD)
  dplyr::filter(vpd>0)
```

# Prepare data for training

Prepare training data, removing NAs and outliers.
```{r}
# ## use all observational soil moisture data
# varnams_soilm <- df_fluxnet %>% 
#   dplyr::select( starts_with("SWC_") ) %>% 
#   dplyr::select( -ends_with("QC") ) %>% 
#   names()

## define settings used by multiple functions as a list
settings <- list( 
  target        = "NEE_VUT_REF", 
  predictors    = c("temp","vpd", "ppfd"), 
  varnams_soilm = "wcont_splash",
  nneurons_good = 10,
  nneurons_all  = 10,
  nrep          = 3,
  package       = "nnet"
  )

df_train <- prepare_trainingdata_fvar( df_fluxnet, settings )
```

# Train models and predict `fvar`

Train models for one set of soil moisture input data. In this case it's observational data.
```{r eval = FALSE}
df_nn_soilm_obs <- train_predict_fvar( 
  df_train,
  settings,
  soilm_threshold    = 0.6,
  weights            = NA, 
  verbose = TRUE
  )
```

Do the training again using another set of soil moisture data from a model.
```{r eval=FALSE}
settings$varnams_soilm <- "wcont_swbm"

df_train <- prepare_trainingdata_fvar( df_fluxnet, settings )

df_nn <- train_predict_fvar( 
  df_train,
  settings,
  soilm_threshold    = 0.6, 
  weights            = NA,
  )
```

# Test performance of NN models

Multiple aspects need to be satisfied:

1. NN$_\text{act}$ has no systematic bias related to the level of soil moisture.
```{r}
df_nn_soilm_obs <- df_nn_soilm_obs %>% 
  mutate(bias_act = nn_act - GPP_NT_VUT_REF,
         bias_pot = nn_pot - GPP_NT_VUT_REF,
         soilm_bin = cut(wcont_splash, 10)
         ) 

df_nn_soilm_obs %>% 
  tidyr::pivot_longer(cols = c(bias_act, bias_pot), names_to = "source", values_to = "bias") %>% 
  ggplot(aes(x = soilm_bin, y = bias, fill = source)) +
  geom_boxplot() +
  geom_hline(aes(yintercept = 0.0), linetype = "dotted") +
  labs(title = "Bias vs. soil moisture")
```

This can be tested using a linear regression model of bias vs. soil moisture, testing wether the slope is significantly different from zero.
```{r}
linmod <- lm(bias_act ~ wcont_splash, data = df_nn_soilm_obs)
testsum <- summary(linmod)
slope_mid <- testsum$coefficients["wcont_splash","Estimate"]
slope_se  <- testsum$coefficients["wcont_splash","Std. Error"]
passtest_bias_vs_soilm <- ((slope_mid - slope_se) < 0 && (slope_mid + slope_se) > 0)
print(passtest_bias_vs_soilm)

df_nn_soilm_obs %>% 
  ggplot(aes(x = wcont_splash, y = bias_act)) +
  geom_point() +
  geom_smooth(method = "lm")
```

2. NN$_\text{pot}$ and NN$_\text{act}$ have no bias during moist days.
```{r}
df_nn_soilm_obs %>% 
  tidyr::pivot_longer(cols = c(bias_act, bias_pot), names_to = "source", values_to = "bias") %>% 
  dplyr::filter(moist) %>% 
  ggplot(aes(y = bias, fill = source)) +
  geom_boxplot() +
  geom_hline(aes(yintercept = 0), linetype = "dotted")

df_nn_soilm_obs %>% 
  dplyr::filter(moist) %>%
  summarise(bias_act = mean(bias_act, na.rm = TRUE), bias_pot = mean(bias_pot, na.rm = TRUE))
```

3. NN$_\text{pot}$ and NN$_\text{act}$ have a high *R*$^2$ and low RMSE during moist days.
```{r}
out_modobs <- df_nn_soilm_obs %>% 
  dplyr::filter(moist) %>% 
  rbeni::analyse_modobs2("nn_pot", "nn_act", type = "heat")
out_modobs$gg
```

4. Fit of NN$_\text{act}$ vs. observed (target) values.
```{r}
df_nn_soilm_obs %>% 
  rbeni::analyse_modobs2("nn_act", "GPP_NT_VUT_REF", type = "heat")
```

All above steps are implemented by function `test_performance_fvar()` which returns a list of all ggplot objects and a data frame with evaluation results for each of the three tests.
```{r}
testresults_fvar <- test_performance_fvar(df_nn_soilm_obs, settings)
```

# Get optimal soil moisture threshold

The two steps above (train/predict and performance evaluation) are executed for a set of soil moisture thresholds. The optimal threshold is then selected based on the different performance criteria.
```{r eval=FALSE}
df_profile_performance <- profile_soilmthreshold_fvar(
  df_train,
  settings,
  weights = NA,
  len = 4
  )
```

Select best threshold as described in Stocker et al. (2018). Their procedure was as follows:

1. Determine the difference in the bias of NN$_\text{pot}$ between dry and moist days and select the three (originally used five) thresholds with the greatest difference.
2. Determine RMSE of NNpot vs. NNact during moist days and take the threshold where it's minimised
```{r}
purrr::map(df_profile_performance, "df_metrics") %>% 
  bind_rows(.id = "threshold") %>% 
  arrange(-diff_ratio_dry_moist) %>% 
  slice(1:3) %>% 
  arrange(rmse_nnpot_nnact_moist) %>% 
  slice(1) %>% 
  pull(threshold) %>% 
  as.numeric()
```

This is implemented by a function:
```{r}
get_opt_threshold(df_profile_performance)
```

Repeat fvar algorithm with optimal threshold.
```{r}
df_nn <- train_predict_fvar( 
  df_train,
  settings,
  soilm_threshold = get_opt_threshold(df_profile_performance),
  weights         = NA
  )
```

<!-- I tried a different approach, too: -->

<!-- Select best threshold, based on the different criteria.  -->
<!-- ```{r} -->
<!-- calc_inverse <- function(x){1/x} -->
<!-- calc_inverse_absdiff_to_one <- function(x){1/abs(x-1)} -->

<!-- df_profile_summary <- df_profile_performance %>%  -->

<!--   ## small bias -->
<!--   mutate_at(vars(starts_with("bias")), list(abs)) %>%  -->

<!--   ## small value is good -> take the inverse -->
<!--   mutate_at(vars(ends_with("rmse")), list(calc_inverse)) %>%  -->

<!--   ## small difference to one is good -->
<!--   mutate_at(vars(ends_with("slope")), list(calc_inverse_absdiff_to_one)) %>%  -->
<!--   mutate_at(vars(starts_with("slope")), list(calc_inverse_absdiff_to_one)) %>%  -->

<!--   ## center and scale all  -->
<!--   mutate_at(c(1, 3:10), list(scale)) %>%  -->
<!--   t() -->

<!-- df_profile_summary <- df_profile_summary %>%  -->
<!--   as_tibble() %>%  -->
<!--   setNames(as.character(paste0("thresh_", df_profile_summary["threshold",]))) %>%  -->
<!--   filter_all(all_vars(!is.nan(.))) %>%  -->
<!--   summarise_all(sum, nan.rm=TRUE) -->
<!-- ``` -->


# Identify soil moisture droughts

Optional: aggregate fvar derived from different soil moisture datasets
```{r eval = FALSE}
df_nn_multiplesoilmsources <- bind_rows( df_nn_soilm_obs, df_nn_soilm_obs ) %>% 
  group_by( date ) %>%
  summarise( 
    fvar       = mean( fvar, na.rm=TRUE ), 
    fvar_min   = min(  fvar, na.rm=TRUE ),
    fvar_max   = max(  fvar, na.rm=TRUE ), 
    fvar_med   = median(  fvar, na.rm=TRUE ), 
    nn_pot = mean( nn_pot, na.rm=TRUE ), 
    nn_act = mean( nn_act, na.rm=TRUE )
    ) %>%
  left_join( select_( df_fluxnet, "date", settings$target ), by = "date" )
```


```{r}
out_droughts <- get_droughts_fvar( 
  df_nn, 
  nam_target     = settings$target, 
  leng_threshold = 10, 
  df_soilm       = select(df_fluxnet, date, soilm=wcont_swbm),
  df_par         = select(df_fluxnet, date, par=ppfd),
  par_runmed     = 10
  )
```

Plot what we got.
```{r}
ggplot() +
  geom_rect(
    data=out_droughts$droughts, 
    aes(xmin=date_start, xmax=date_end, ymin=-99, ymax=99), 
    fill=rgb(0,0,0,0.3), 
    color=NA) + 
  geom_line(data=out_droughts$df_nn, aes(date, fvar_smooth_filled)) +
  coord_cartesian(ylim=c(0, 1.2))
```

# Align data by droughts

```{r}
dovars <- c("GPP_NT_VUT_REF", "fvar", "soilm", "nn_pot", "nn_act")
df_nn <- out_droughts$df_nn %>% mutate(site="FR-Pue")
df_alg <- align_events( 
  select( df_nn, site, date, one_of(dovars)), 
  select( df_nn, site, date, isevent=is_drought_byvar_recalc ),
  dovars,
  do_norm=FALSE,
  leng_threshold=10, 
  before=20, after=80, nbins=10
  )
```


Plot what we got
```{r}
df_alg$df_dday_agg_inst_site %>% 
  ggplot(aes(dday, fvar_median)) +
  geom_ribbon(aes(ymin=fvar_q33, ymax=fvar_q66), fill="grey70") +
  geom_line() +
  geom_vline(xintercept=0, linetype="dotted") +
  ylim(0,1) + 
  labs(title = "fVAR")

median <- df_alg$df_dday_agg_inst_site %>%
  select(dday, nn_pot=nn_pot_median, nn_act=nn_act_median, gpp_obs=GPP_NT_VUT_REF_median) %>% 
  tidyr::gather(model, var_nn_median, c(gpp_obs, nn_pot, nn_act))

q33 <- df_alg$df_dday_agg_inst_site %>%
  select(dday, nn_pot=nn_pot_q33, nn_act=nn_act_q33, gpp_obs=GPP_NT_VUT_REF_q33) %>% 
  tidyr::gather(model, var_nn_q33, c(gpp_obs, nn_pot, nn_act))

q66 <- df_alg$df_dday_agg_inst_site %>%
  select(dday, nn_pot=nn_pot_q66, nn_act=nn_act_q66, gpp_obs=GPP_NT_VUT_REF_q66) %>% 
  tidyr::gather(model, var_nn_q66, c(gpp_obs, nn_pot, nn_act))

df_tmp <- median %>% 
  left_join(q33, by=c("dday", "model")) %>% 
  left_join(q66, by=c("dday", "model"))

df_tmp %>% 
  ggplot(aes(dday, var_nn_median, color=model)) +
  geom_line() +
  geom_ribbon(aes(ymin=var_nn_q33, ymax=var_nn_q66, fill=model), alpha=0.3, color=NA) +
  geom_vline(xintercept=0, linetype="dotted") +
  labs(
    title = "GPP", 
    subtitle = "Observation and models",
    x="dday", y="GPP (g C m-2 d-1)") +
  scale_color_discrete(
    name="Model", 
    breaks=c("transp_obs", "nn_act", "nn_pot"),
    labels=c("Observed", expression(NN[act]), expression(NN[pot])) ) +
  scale_fill_discrete(
    name="Model", 
    breaks=c("transp_obs", "nn_act", "nn_pot"),
    labels=c("Observed", expression(NN[act]), expression(NN[pot])) )
```
