---
title: "fvar with soil respiration data"
author: "Beni Stocker"
date: "4/30/2019"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
library(rbeni)
library(dplyr)
library(ggplot2)
library(readr)
library(lubridate)
library(tidyr)

source("get_obs_bysite_fluxnet2015.R")
source("train_predict_fvar.R")
source("predict_nn.R")
source("prepare_trainingdata_fvar.R")
source("remove_outliers.R")
source("get_consecutive.R")
source("add_alpha.R")
source("align_events.R")
source("get_droughts_fvar.R")
```

## Load data

```{r}
hdf <- read_csv("/alphadata01/bstocker/data/resp_soil_avni/moflux2015_senttoBeni_bs.csv") %>% 
  
  ## get clean date
  # slice(1:100) %>% 
  # dplyr::select(1:13) %>% 
  rowwise() %>% 
  mutate(date = ymd_hm(paste0(as.character(Year), "-01-01 00:00")) + days( DOY - 1 ) + minutes( HourMin ) )
save(hdf, file = "./data/hdf.Rdata")
```

Gather data from multiple chambers (assuming they sample the same system).
```{r}
hdf_g_co2flux <- hdf %>% 
  select(date, ends_with("co2flux")) %>% 
  tidyr::gather(chamber, co2flux, ends_with("co2flux")) %>% 
  mutate(chamber = str_remove(chamber, "co2flux"))

hdf_g_soilm <- hdf %>% 
  select(date, ends_with("soilm")) %>% 
  tidyr::gather(chamber, soilm, ends_with("soilm")) %>% 
  mutate(chamber = str_remove(chamber, "soilm"))

hdf_g_soilt <- hdf %>% 
  select(date, ends_with("soilt")) %>% 
  tidyr::gather(chamber, soilt, ends_with("soilt")) %>% 
  mutate(chamber = str_remove(chamber, "soilt"))

hdf_g <- hdf_g_co2flux %>% 
  left_join(hdf_g_soilm, by = c("date", "chamber")) %>% 
  left_join(hdf_g_soilt, by = c("date", "chamber")) %>% 
  left_join(select(hdf, date, ORNL72NEECO2), by = "date")
```

Fig GAM 
```{r}
library(mgcv)
library(mgcViz)

mod_gam <- gam( co2flux ~ s(soilt) + s(soilm), data = hdf_g, method = "REML" )
mod_gam <- getViz(mod_gam)

gg_soilt <- plot( sm(mod_gam, 1) )
print(gg_soilt)

gg_soilm <- plot( sm(mod_gam, 2) )
print(gg_soilm)

gg_soilt + 
  l_fitLine(colour = "red") + 
  # l_rug(mapping = aes(x=x, y=y), alpha = 0.8) +
  l_ciLine(mul = 5, colour = "red", linetype = 2) + 
  l_points(shape = 19, size = 1, alpha = 0.1) + 
  theme_classic() +
  xlim(0, 0.5) +
  stat_smooth(method = "lm", col = "blue") +
  labs(x = "soilt", y = "co2flux")

gg_soilm + 
  l_fitLine(colour = "red") + 
  # l_rug(mapping = aes(x=x, y=y), alpha = 0.8) +
  l_ciLine(mul = 5, colour = "red", linetype = 2) + 
  l_points(shape = 19, size = 1, alpha = 0.1) + 
  theme_classic() +
  xlim(0, 0.5) +
  stat_smooth(method = "lm", col = "blue") +
  labs(x = "soilm", y = "co2flux")
```


# Prepare data for training

Prepare training data, removing NAs and outliers.
```{r}
## use all observational soil moisture data
varnams_soilm <- hdf_g %>% 
  dplyr::select( ends_with("soilm") ) %>% 
  names()

## define the target variable and predictors
settings <- list( 
  target        = "co2flux", 
  predictors    = c("soilm", "soilt", "ORNL72NEECO2"), 
  varnams_soilm = varnams_soilm 
  )

df_train <- hdf_g %>% 
  dplyr::mutate_at( vars(settings$target), funs( remove_outliers(.) ) ) %>% 
  dplyr::filter( soilt > 5.0 ) %>%
  dplyr::filter( yday(date) > 90 ) %>% 
  tidyr::drop_na()
```

# some plotting

```{r}
df_train %>% 
  ggplot() +
  geom_line(aes(x = date, y = co2flux, color = chamber))

df_train %>% 
  ggplot() +
  geom_point(aes(x = date, y = soilm, color = chamber))

df_train %>% 
  ggplot() +
  geom_histogram(aes(soilm))
```

# Set a soil moisture threshold

To be done properly ... later.
```{r}
soilm_threshold <- 0.3
```

# Train models and predict `fvar`

Train models for one set of soil moisture input data. In this case it's observational data.
```{r}
df_nn_soilm_obs <- train_predict_fvar( 
  df_train,
  settings,
  soilm_threshold    = soilm_threshold, # hard coded here instead of using output from profile_soilmthreshold_fvar()
  hidden_good        = 10,   # hard coded here instead of using output from profile_soilmthreshold_fvar()
  hidden_all         = 10,   # hard coded here instead of using output from profile_soilmthreshold_fvar()
  nrep               = 1, 
  weights            = NA, 
  package            = "nnet" 
  )
```

# Test performance of NN models

```{r}
df_modobs <- df_nn_soilm_obs %>% 
  left_join(select(df_train, date, settings$target), by="date") %>% 
  left_join(select(df_train, date, soilm, soilt), by = "date") %>% 
  mutate( moist = df_nn_soilm_obs$moist )

analyse_modobs2(df_modobs, "var_nn_act", "co2flux")
analyse_modobs2(df_modobs, "var_nn_pot", "co2flux")

ggplot(df_modobs)+
  geom_point(aes(x=var_nn_pot, y=co2flux, color=moist))
```

# Functional relationship

```{r}
df_modobs %>% 
  ggplot() +
  geom_point(aes(x = soilm, y = fvar))

df_modobs %>% 
  ggplot() +
  geom_point(aes(x = soilt, y = fvar))
```


