---
title: "fvar R package"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r}
devtools::load_all(".")
library(rsofun)
library(dplyr)
library(readr)
library(lubridate)
library(ggplot2)
```


# Get data

Training data is to be read in as a data frame with column names corresponding to the variable names defined by `settings` (see below).

Read observational data including soil moisture for one site (`"AU-How"`) from a FLUXNET 2015 data file (standard). In the present example
```{r}
load("./data/df_fluxnet.Rdata")
```

# Prepare data for training

Prepare training data, removing NAs and outliers.
```{r}
## use all observational soil moisture data
varnams_soilm <- df_fluxnet %>% 
  dplyr::select( starts_with("SWC_") ) %>% 
  dplyr::select( -ends_with("QC") ) %>% 
  names()

## define settings used by multiple functions as a list
settings <- list( 
  target        = "GPP_NT_VUT_REF", 
  predictors    = c("temp_day","vpd_day", "ppfd", "fpar"), 
  varnams_soilm = varnams_soilm,
  nneurons_good = 10,
  nneurons_all  = 10,
  nrep          = 3,
  package       = "nnet"
  )

df_train <- prepare_trainingdata_fvar( df_fluxnet, settings )
```

# Train models and predict `fvar`

Train models for one set of soil moisture input data. In this case it's observational data.
```{r eval = FALSE}
df_nn_soilm_obs <- train_predict_fvar( 
  df_train,
  settings,
  soilm_threshold    = 0.6, # hard coded here instead of using output from profile_soilmthreshold_fvar()
  weights            = NA, 
  verbose = TRUE
  )
```

Do the training again using another set of soil moisture data from a model.
```{r}
settings$varnams_soilm <- "wcont_swbm"

df_train <- prepare_trainingdata_fvar( df_fluxnet, settings )

df_nn_soilm_swbm <- train_predict_fvar( 
  df_train,
  settings,
  soilm_threshold    = 0.6, 
  weights            = NA,
  )
```

# Test performance of NN models

Multiple aspects need to be satisfied:

1. NN$_\text{act}$ has no systematic bias related to the level of soil moisture.
```{r}
df_nn_soilm_swbm <- df_nn_soilm_swbm %>% 
  mutate(bias_act = nn_act_vals - GPP_NT_VUT_REF,
         bias_pot = nn_pot_vals - GPP_NT_VUT_REF,
         soilm_bin = cut(wcont_swbm, 10)
         ) 

df_nn_soilm_swbm %>% 
  tidyr::pivot_longer(cols = c(bias_act, bias_pot), names_to = "source", values_to = "bias") %>% 
  ggplot(aes(x = soilm_bin, y = bias, fill = source)) +
  geom_boxplot() +
  geom_hline(aes(yintercept = 0.0), linetype = "dotted") +
  labs(title = "Bias vs. soil moisture")
```

This can be tested using a linear regression model of bias vs. soil moisture, testing wether the slope is significantly different from zero.
```{r}
linmod <- lm(bias_act ~ wcont_swbm, data = df_nn_soilm_swbm)
testsum <- summary(linmod)
slope_mid <- testsum$coefficients["wcont_swbm","Estimate"]
slope_se  <- testsum$coefficients["wcont_swbm","Std. Error"]
passtest_bias_vs_soilm <- ((slope_mid - slope_se) < 0 && (slope_mid + slope_se) > 0)
print(passtest_bias_vs_soilm)

df_nn_soilm_swbm %>% 
  ggplot(aes(x = wcont_swbm, y = bias_act)) +
  geom_point() +
  geom_smooth(method = "lm")
```

2. NN$_\text{pot}$ and NN$_\text{act}$ have no bias during moist days.
```{r}
df_nn_soilm_swbm %>% 
  tidyr::pivot_longer(cols = c(bias_act, bias_pot), names_to = "source", values_to = "bias") %>% 
  dplyr::filter(moist) %>% 
  ggplot(aes(y = bias, fill = source)) +
  geom_boxplot() +
  geom_hline(aes(yintercept = 0), linetype = "dotted")

df_nn_soilm_swbm %>% 
  dplyr::filter(moist) %>%
  summarise(bias_act = mean(bias_act, na.rm = TRUE), bias_pot = mean(bias_pot, na.rm = TRUE))
```

3. NN$_\text{pot}$ and NN$_\text{act}$ have a high *R*$^2$ and low RMSE during moist days.
```{r}
out_modobs <- df_nn_soilm_swbm %>% 
  dplyr::filter(moist) %>% 
  rbeni::analyse_modobs2("nn_pot_vals", "nn_act_vals", type = "heat")
out_modobs$gg
```

4. Fit of NN$_\text{act}$ vs. observed (target) values.
```{r}
df_nn_soilm_swbm %>% rbeni::analyse_modobs2("nn_act_vals", "GPP_NT_VUT_REF", type = "heat")
```

All above steps are implemented by function `test_performance_fvar()` which returns a list of all ggplot objects and a data frame with evaluation results for each of the three tests.
```{r}
testresults_fvar <- test_performance_fvar(df_nn_soilm_swbm)
```

# Get optimal soil moisture threshold

The two steps above (train/predict and performance evaluation) are executed for a set of soil moisture thresholds. The optimal threshold is then selected based on the different performance criteria.
```{r eval=FALSE}
df_profile_performance <- profile_soilmthreshold_fvar(
  df_train,
  settings,
  weights = NA,
  len = 4
  )
```

Select best threshold, based on the different criteria. 
```{r}
calc_inverse <- function(x){1/x}
calc_inverse_absdiff_to_one <- function(x){1/abs(x-1)}

df_profile_summary <- df_profile_performance %>% 
  
  ## small bias
  mutate_at(vars(starts_with("bias")), list(abs)) %>% 
  
  ## small value is good -> take the inverse
  mutate_at(vars(ends_with("rmse")), list(calc_inverse)) %>% 

  ## small difference to one is good
  mutate_at(vars(ends_with("slope")), list(calc_inverse_absdiff_to_one)) %>% 
  mutate_at(vars(starts_with("slope")), list(calc_inverse_absdiff_to_one)) %>% 
  
  ## center and scale all 
  mutate_at(c(1, 3:10), list(scale)) %>% 
  t()

df_profile_summary <- df_profile_summary %>% 
  as_tibble() %>% 
  setNames(as.character(paste0("thresh_", df_profile_summary["threshold",]))) %>% 
  filter_all(all_vars(!is.nan(.))) %>% 
  summarise_all(sum, nan.rm=TRUE)
```


# Identify soil moisture droughts

```{r}
## first aggregate fvar derived from different soil moisture datasets
df_nn <- bind_rows( df_nn_soilm_obs, df_nn_soilm_swbm ) %>% 
  group_by( date ) %>%
  summarise( 
    fvar       = mean( fvar, na.rm=TRUE ), 
    fvar_min   = min(  fvar, na.rm=TRUE ),
    fvar_max   = max(  fvar, na.rm=TRUE ), 
    fvar_med   = median(  fvar, na.rm=TRUE ), 
    var_nn_pot = mean( var_nn_pot, na.rm=TRUE ), 
    var_nn_act = mean( var_nn_act, na.rm=TRUE )
    ) %>%
  left_join( select_( df_fluxnet, "date", settings$target ), by = "date" )

out_droughts <- get_droughts_fvar( 
  df_nn, 
  nam_target     = settings$target, 
  leng_threshold = 10, 
  df_soilm       = select(df_fluxnet, date, soilm=wcont_swbm),
  df_par         = select(df_fluxnet, date, par=ppfd),
  par_runmed     = 10
  )
```

Plot what we got.
```{r}
ggplot() +
  geom_rect(
    data=out_droughts$droughts, 
    aes(xmin=date_start, xmax=date_end, ymin=-99, ymax=99), 
    fill=rgb(0,0,0,0.3), 
    color=NA) + 
  geom_line(data=out_droughts$df_nn, aes(date, fvar_smooth_filled)) +
  coord_cartesian(ylim=c(0, 1))
```

# Align data by droughts

```{r}
dovars <- c("transp_obs", "fvar", "soilm", "var_nn_pot", "var_nn_act")
df_nn <- out_droughts$df_nn %>% mutate(site="AU-How")
df_alg <- align_events( 
  select( df_nn, site, date, one_of(dovars)), 
  select( df_nn, site, date, isevent=is_drought_byvar_recalc ),
  dovars,
  do_norm=FALSE,
  leng_threshold=10, 
  before=20, after=80, nbins=10
  )
```


Plot what we got
```{r}
df_alg$df_dday_agg_inst_site %>% 
  ggplot(aes(dday, fvar_median)) +
  geom_ribbon(aes(ymin=fvar_q33, ymax=fvar_q66), fill="grey70") +
  geom_line() +
  geom_vline(xintercept=0, linetype="dotted") +
  ylim(0,1) + 
  labs(title = "fVAR")

median <- df_alg$df_dday_agg_inst_site %>%
  select(dday, var_nn_pot=var_nn_pot_median, var_nn_act=var_nn_act_median, transp_obs=transp_obs_median) %>% 
  tidyr::gather(model, var_nn_median, c(transp_obs, var_nn_pot, var_nn_act))

q33 <- df_alg$df_dday_agg_inst_site %>%
  select(dday, var_nn_pot=var_nn_pot_q33, var_nn_act=var_nn_act_q33, transp_obs=transp_obs_q33) %>% 
  tidyr::gather(model, var_nn_q33, c(transp_obs, var_nn_pot, var_nn_act))

q66 <- df_alg$df_dday_agg_inst_site %>%
  select(dday, var_nn_pot=var_nn_pot_q66, var_nn_act=var_nn_act_q66, transp_obs=transp_obs_q66) %>% 
  tidyr::gather(model, var_nn_q66, c(transp_obs, var_nn_pot, var_nn_act))

df_tmp <- median %>% 
  left_join(q33, by=c("dday", "model")) %>% 
  left_join(q66, by=c("dday", "model"))

df_tmp %>% 
  ggplot(aes(dday, var_nn_median, color=model)) +
  geom_line() +
  geom_ribbon(aes(ymin=var_nn_q33, ymax=var_nn_q66, fill=model), alpha=0.3, color=NA) +
  geom_vline(xintercept=0, linetype="dotted") +
  labs(
    title = "Transpiration", 
    subtitle = "Observation and models",
    x="dday", y="Transpiration (XXX)") +
  scale_color_discrete(
    name="Model", 
    breaks=c("transp_obs", "var_nn_act", "var_nn_pot"),
    labels=c("Observed", expression(NN[act]), expression(NN[pot])) ) +
  scale_fill_discrete(
    name="Model", 
    breaks=c("transp_obs", "var_nn_act", "var_nn_pot"),
    labels=c("Observed", expression(NN[act]), expression(NN[pot])) )
```
